---
title: "SAMsize"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---


#Setup

## Normalize path and load libraries etc.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=normalizePath('../../'))
```


```{r,message=FALSE, warning=FALSE}
library(gsl)
library(dplyr)
library(magrittr)
library(cowplot)
library(viridis)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cols <- viridis(10)

## more packages
library(nlme)
library(quantgen)
library(tidyr)
```

### Functions
```{r}
countsize <- read.csv("data/SAM_cellcount.csv", header=T)
head(countsize)
```

In the above SAM Cell count table, `r length(unique(countsize$Genotype))` Genotypes were collected for SAM cell counts in `r length(table(countsize$Growth_Period))` growth periods, each period with 3 plants. 
There are also two factors, BAK1 and SDA1, that associated with some traits. 
From the data, we learned that `Count_Cells` is significantly correlated with `SAM_V` ($r=0.75, P value < 0.01$). But `Count_Cells` is not correlated with `Mean_Area_percell` ($r=0.03, P value = 0.7$).

We estimated BLUE values by fitting a linear mixed model, where `Genotype` as a fixed effect and `Plant` as a random effect.


```{r}
csg1 <- mixed_model(data = subset(countsize, Growth_Period %in% "G1"),
                    model = Count_Cells ~ Genotype, random = ~1 | Plant)

csg2 <- mixed_model(data = subset(countsize, Growth_Period %in% "G2"),
                    model = Count_Cells ~ Genotype, random = ~1 | Plant)

csg3 <- mixed_model(data = subset(countsize, Growth_Period %in% "G3"),
                    model = Count_Cells ~ Genotype, random = ~1 | Plant, trait = "CS_G3")

cs <- merge(csg1, csg2, by="Genotype")
cs <- merge(cs, csg3, by="Genotype") %>% set_names(c("genotype", "g1", "g2", "g3"))
lcs <- cs %>% gather(key="Growth", value="cellnum", 2:4)

#theme_set(theme_grey(base_size = 18)) 
p <- ggplot(lcs, aes(x=toupper(Growth), y=cellnum, fill = Growth)) + 
    #opts(axis.text.x=theme_text(angle=90)) +
    theme_bw(base_size = 18) +
    #theme(axis.text.x = element_text(angle = 90, hjust = 1, size=12)) +
    geom_boxplot() +
    ggtitle("BLUEs of three growth periods") + xlab("") + ylab("Cell Number") + 
    #ggtitle("Sequencing Depth") + xlab("") + ylab("Depth per cytosine site") + 
    guides(fill=FALSE)
p

```

```{r, eval=T}
#samsize<-read.csv("~/Desktop/samsize.csv",header=T)
#samsize_unsummary<-read.csv("~/Desktop/samsize_unsummarized.csv",header=T)
plantstuff <- read.csv("data/plantstuff.csv",header=T) %>%
  select(Genotype,PlantH..cm.,EarH..cm.,mean_nodes,mean_dia_ear..mm.,mean_dia_below..mm.,DTA,SAM_V..um.3.) %>% 
  set_names(c("genotype","plant_height","ear_height","leaf_nodes","ear_width","stem_width","DTA","SAM_volume")) 
plantstuff$genotype <- toupper(plantstuff$genotype)
cs$genotype <- toupper(cs$genotype)

plantcount <- merge(plantstuff, cs, by="genotype")

head(plantcount)
plot(plantcount$DTA, plantcount$g1)
```



```{r, warning=FALSE}
lcs <- cs %>% gather(key="Growth", value="cellnum", 2:4)

p1 <- plantcount[, c("genotype", "g1", "g2", "g3")] %>% 
  gather(key="Growth", value="cellnum", 2:4) %>%
  merge(plantcount[, 1:8], by="genotype")
#DTA, cell count
ggplot(p1, aes(y=DTA, x=cellnum))+
  geom_point(aes(color=genotype))+
  geom_smooth(color="dark grey")+
  facet_wrap(~Growth, scales="free_x")+
  xlab("Cell Number")+
  ylab("Flowering Time")

summary(lm(data=p1, DTA ~ cellnum + Growth))
```


### Fit Linear Mixed Model with relatedness matrix as random

```{r}
fam0 <- read.table("cache/GBSv2.7_id14_flt.fam", header=F)

idcurated <- read.csv("cache/cellnum_GBS_sampleid_curated.csv")
fam <- merge(fam0, idcurated[, c("FullName", "DNASample")], by.x="V1", by.y="FullName", sort=FALSE)
fam$DNASample <- toupper(fam$DNASample)

#### Relatedness estimated from GEMMA
mx <- read.table("cache/relate.cXX.txt")
heatmap(mx)
row.names(mx) <- fam$DNASample
names(mx) <- fam$DNASample

library(d3heatmap)
d3heatmap(mx, scale = "column", dendrogram = "none", color = "Blues")
```






```{r, eval=FALSE}
#cell count, SAM volume
ggplot(plantcount,aes(y=SAM_volume,x=cell_number))+
  geom_point(color=cols[1])+
  geom_smooth(method="loess",color="dark grey")+
  facet_wrap(~growth_period,scales="free_x")+
  ylab("SAM volume")+
  xlab("Cell Number")

#cell size, SAM volume
ggplot(plantcount,aes(y=SAM_volume,x=cell_size))+
  geom_point(color=cols[2])+
  geom_smooth(method="loess",color="dark grey")+
  facet_wrap(~growth_period,scales="free_x")+
  ylab("SAM volume")+
  xlab("Cell Size")

#cell size, cell number
ggplot(plantcount,aes(x=cell_number,y=cell_size))+
  geom_point(color=cols[3])+
  geom_smooth(method="loess",color="dark grey")+
  facet_wrap(~growth_period,scales="free_x")+
  xlab("Cell Number")+
  ylab("Cell Size")



#more cells -> no diff plant height. smaller ear height, weakly smaller ear width & stem_width, shorter DTA

summary(lm(data=plantcount,DTA~cell_number+SDA1+BAK1))
```




